
\subsection*{7.2 Simple and Compound Interest}

Interest is the extra amount paid on a loan or earned on savings over time. There are two main types of interest calculations: simple interest and compound interest.

\textbf{Key Concepts:}
\begin{itemize}
	\item \textbf{Simple Interest:} Interest calculated on the original principal amount.
	\[
	I = P \times r \times t
	\]
	where:
	\begin{itemize}
		\item $I$ = Interest
		\item $P$ = Principal (initial amount)
		\item $r$ = Annual interest rate (as a decimal)
		\item $t$ = Time in years
	\end{itemize}
	
	\item \textbf{Total Amount with Simple Interest:}
	\[
	A = P + I = P(1 + rt)
	\]
	
	\item \textbf{Compound Interest:} Interest calculated on both the principal and previously earned interest.
	\[
	A = P \left(1 + \frac{r}{n} \right)^{nt}
	\]
	where:
	\begin{itemize}
		\item $A$ = Final amount after interest
		\item $P$ = Principal
		\item $r$ = Annual interest rate (as a decimal)
		\item $n$ = Number of times interest is applied per year
		\item $t$ = Time in years
	\end{itemize}
	
	\item \textbf{Continuous Compound Interest:}
	\[
	A = P e^{rt}
	\]
	where $e$ is the mathematical constant $\approx 2.718$.
\end{itemize}

\textbf{Examples:}

\begin{flushleft}
	\textbf{Example 1: Simple Interest Calculation}
	
	John deposits \$5000 in a bank account that pays 4\% simple interest per year. Find the total interest earned after 3 years.
	
	\textbf{Solution:}
	
	Step 1: Use the simple interest formula.
	\[
	I = 5000 \times 0.04 \times 3
	\]
	\[
	I = 600
	\]
	
	Step 2: Find the total amount.
	\[
	A = P + I = 5000 + 600 = 5600
	\]
	
	Thus, John earns **\$600** in interest, and his total amount after 3 years is **\$5600**.
\end{flushleft}

\begin{flushleft}
	\textbf{Example 2: Compound Interest Calculation}
	
	Sarah invests \$2000 in a savings account that offers an annual interest rate of 5\%, compounded yearly. Find the total amount after 4 years.
	
	\textbf{Solution:}
	
	Step 1: Use the compound interest formula.
	\[
	A = 2000 \left(1 + \frac{0.05}{1} \right)^{1 \times 4}
	\]
	\[
	A = 2000 \left(1.05 \right)^4
	\]
	\[
	A = 2000 \times 1.21550625
	\]
	\[
	A = 2431.01
	\]
	
	Thus, Sarah's total amount after 4 years is **\$2431.01**.
\end{flushleft}