
\subsection*{7.3 Depreciation and Appreciation}

Depreciation and appreciation refer to the decrease or increase in the value of an asset over time.

\textbf{Key Concepts:}
\begin{itemize}
	\item \textbf{Depreciation:} The reduction in the value of an asset over time, often due to wear and tear or obsolescence.
	\[
	A = P(1 - r)^t
	\]
	where:
	\begin{itemize}
		\item $A$ = Value of the asset after time $t$
		\item $P$ = Initial value (purchase price)
		\item $r$ = Rate of depreciation (as a decimal)
		\item $t$ = Time in years
	\end{itemize}
	
	\item \textbf{Appreciation:} The increase in the value of an asset over time, often due to demand or inflation.
	\[
	A = P(1 + r)^t
	\]
	where:
	\begin{itemize}
		\item $A$ = Value of the asset after time $t$
		\item $P$ = Initial value
		\item $r$ = Rate of appreciation (as a decimal)
		\item $t$ = Time in years
	\end{itemize}
\end{itemize}

\textbf{Examples:}

\begin{flushleft}
	\textbf{Example 1: Depreciation Calculation}
	
	A car was bought for \$15,000 and depreciates at a rate of 10\% per year. Find its value after 5 years.
	
	\textbf{Solution:}
	
	Step 1: Use the depreciation formula.
	\[
	A = 15000 (1 - 0.10)^5
	\]
	\[
	A = 15000 (0.9)^5
	\]
	\[
	A = 15000 \times 0.59049
	\]
	\[
	A = 8857.35
	\]
	
	Thus, the car's value after 5 years is **\$8857.35**.
\end{flushleft}

\begin{flushleft}
	\textbf{Example 2: Appreciation Calculation}
	
	A house was purchased for \$120,000 and appreciates at a rate of 5\% per year. Find its value after 8 years.
	
	\textbf{Solution:}
	
	Step 1: Use the appreciation formula.
	\[
	A = 120000 (1 + 0.05)^8
	\]
	\[
	A = 120000 (1.477455)
	\]
	\[
	A = 177294.60
	\]
	
	Thus, the house's value after 8 years is **\$177,294.60**.
\end{flushleft}

